---
title: "MSIA 400 - Assignment 3"
output: html_document
---
# MSIA 400 - Assignment 3
## *Steven Lin*


## Setup

```{r Setup, results='hide'}

# Setup ####

# My PC
main = "/Users/Steven/Documents/Academics/3_Graduate School/2014-2015_NU/"

# Aginity
#main = "\\\\nas1/labuser169"

course = "MSIA_400_Analytics for Competitive Advantage"
datafolder = "Lab/Assignment_03"
setwd(file.path(main,course, datafolder))

opts_knit$set(root.dir = getwd()) # set the working directory for all chunks

```

## Problem 1

```{r P1, results='hide'}
# Import data
filename = "markov100.txt"
mydata= read.table(filename, header=F, 
                   col.names=paste("S", 1:100, sep = ""),
                   row.names=paste("S", 1:100, sep = ""))
P = as.matrix(mydata)

# Look at data
names(mydata)
head(mydata)[,1:6]
nrow(mydata)
#summary(mydata)

n = dim(P)[1]

```

# Part a
```{r }
# install.packages("expm")
library(expm)
end = 5
a = rep(0,n)
a[1]=1
ntrans = 10
dist = a %*% (P %^% ntrans)
dist[end]

```

Probability of being in State 5 after 10 transitions: **`r dist[end]`**

# Part b
```{r }
# install.packages("expm")
library(expm)
end = 10
a = rep(0,n)
a[c(1,2,3)]=1/3
ntrans = 10
dist = a %*% (P %^% ntrans)
dist[end]

```

Suppose we are at one of States 1,2, and 3 with equal probabilities. The probability of being in State 10 after 10 transitions : **`r dist[end]`**

# Part c
```{r }
# install.packages("expm")
library(expm)

end = 1
Q = t(P)- diag(n)
Q[n,] = rep(1,n)
rhs = rep(0,n); rhs[n]=1
Pi = solve(Q) %*% rhs
Pi[end]

```

Steady state probability of being in State 1: **`r Pi[end]`**

# Part d
```{r }
# install.packages("expm")
library(expm)


# Mean first passage time m ij is the expected number of transitions before
# we first reach state j, given we are currently in i

# Mean first passage time to state Start
# B: submatrix of P obtained by deleting the row and column corresponding to state Sart
# m: vector of mij, i diff Start, j = start
# e: vector of 1's 
# m = e + Bm
# m = (I-B)^-1e

start = 1
end = 100

B =P[-end,-end]
Q = diag(n-1) - B
e = rep(1,n-1)
m = solve(Q) %*% e
m[paste("S", start, sep = ""),]
```

Mean first passage time from State 1 to State 100: **`r m[paste("S", start, sep = ""),]`**

## Problem 2

```{r P2, results='hide'}
# Import data
filename = "webtraffic.txt"
mydata= read.table(filename, header=T)
P = as.matrix(mydata)

# Look at data
names(mydata)
head(mydata)[,1:6]
nrow(mydata)
#summary(mydata)

```
# Part a

```{r }
# sum columns (traffic from tij column) and store in traffic matrix
# counts total traffic between state i to state j
Traffic = matrix(as.matrix(colSums(P)),ncol=9,byrow=T)
Traffic 

```
# Part b
```{r }
Traffic[9,1]=1000 #why 9,1 and not 9,9?

sumRows = rowSums(Traffic)
P = Traffic

for (i in 1:9){
  P[i,]=P[i,]/sumRows[i]
}
colnames(P)=paste("p",1:9,sep="")
rownames(P)=paste("p",1:9,sep="")

# the one step transition probability matrix
P
```

# Part c
```{r }

library(expm)

n = dim(P)[1]
Q = t(P)- diag(n)
Q[n,] = rep(1,n)
rhs = rep(0,n); rhs[n]=1
Pi = solve(Q) %*% rhs

# steady state probability vector
Pi



```

# Part d
```{r }

avgTimePage = c(0.1,2,3,5,5,3,3,2,0) # add zero time p9
names(avgTimePage) = paste("p",1:9,sep="")

# this is like an upperbound, know that the average time spent in system has to be lower than this
sum(avgTimePage )

# p i is the fraction of time the process is in state i (in the long-run)
# this means 0.15 of the time, the user is in page 1
# the lon-run average time per period is time p1 * long-run time prob p1 (steady state probability state 1)

avgTimeSite2 = sum(avgTimePage*Pi)
avgTimeSite2

# Assume start state x, expected number of step until the first return is 1/p(x)
1/Pi[1,1]

# Mean first passage time m ij is the expected number of transitions before
# we first reach state j, given we are currently in i

# Mean first passage time to state 9
# B: submatrix of P obtained by deleting the row and column corresponding to state 9
# m: vector of mij, i diff 9, j = 9
# e: vector of 1's 
# m = e + Bm
# m = (I-B)^-1e

start = 1
end = 9

B =P[-end,-end]
Q = diag(n-1) - B
e = rep(1,n-1)
m = solve(Q) %*% e
m

# ave time p1 * # times in p1 + ave time p2 * # times in p2 / (# times in p1 + p2)

# repeat mean first passage time but look at time instead of transitions
# mij = pij * tj + sum(k diff j) pik(tk+mkj)
# mij = e + Bm, where e= pij * tj + sum(k diff j) pik(tk)
# tj = 0 for state 9
# then add ti (othwerwise will double count)

start = 1
end = 9

B =P[-end,-end]
Q = diag(n-1) - B

t = avgTimePage[-end] # there is no time for state 9
e = P[-end,end]*avgTimePage[end]+B%*%t
m = solve(Q) %*% e
m

avgTimeSite = m[paste("p", start, sep = ""),]+t[start]
avgTimeSite 

# Average time on stie = Total time on site/visits
# total time on site = page a + page bb
# average time on page = total time on page a/(pageviews-Exits)

# Other option is to sum all visits to page i, and do sum(avg tim page i * # visits page i)/sum(vists)

sum(apply(Traffic,2,sum)*avgTimePage)/sum(Traffic[,1])

```

The average time a visitor spend on the website: `r avgTimeSite`. This is using the approach of mean passage time. Alternatively, multiplying the average times by the number of visits (total time spent in system) divided by the number of visitors gives the same answer. Using your approach (multiply steady state by average times) the answer is: `r avgTimeSite2`. Note that this number seems to low, since we know most customers visit more than one page and the lowest average time per page is 2 min (without counting page 1). 


# Part e
```{r }

# from Page2, 30% of the current outgoing traffic to State 3 would move to State 6
Traffic[2,6]=Traffic[2,3]*0.3
Traffic[2,3]=Traffic[2,3]-Traffic[2,6]

# 20% of the current outgoing traffic to State 4 would move to State 7
Traffic[2,7]=Traffic[2,4]*0.2
Traffic[2,4]=Traffic[2,4]-Traffic[2,7]

Traffic

sumRows = rowSums(Traffic)
P = Traffic

for (i in 1:9){
  P[i,]=P[i,]/sumRows[i]
}
colnames(P)=paste("p",1:9,sep="")
rownames(P)=paste("p",1:9,sep="")

# the one step transition probability matrix
P

library(expm)

n = dim(P)[1]
Q = t(P)- diag(n)
Q[n,] = rep(1,n)
rhs = rep(0,n); rhs[n]=1
Pi2 = solve(Q) %*% rhs

# steady state probability vector
Pi2

# Decide if the link helped balancing the traffic by comparing the variance of Pi and Pi2.
Pi-Pi2
var(Pi)
var(Pi2)
var(Pi2)<var(Pi)

(var(Pi2)-var(Pi))/var(Pi)


```

The variance of Pi2 is less than the variance of Pi, so adding the links helped reduced the variance and thus balanace the traffic. More specifically, the variance was reduced by `r abs(var(Pi2)-var(Pi))/var(Pi)*100` %. 
